---
title: "Problem Set 2"
output: html_document
date: '2023-10-04'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
set.seed(1)
```

Write a function that performs spectral clustering and call it spec_clut. Save it in a script called spec_clus.R. The function Should:

Take as inputs the normalised Laplacian matrix and the number of communities k.

Return a vector of community indices for the vertices.

Set the nstart argument of k-means to 1000 in your function. Load in the campnet data set up an undirected network for it and test out your function

```{r}

library(igraph)
library(readr)

campnet_Copy <- read.csv("C:/Users/asger/OneDrive - Aarhus universitet/UNI/Machine Learning/Machine_Learning/Problemset 2/Data/campnet - Copy.csv", row.names = 1)
campnet_Attribute <- read.csv("C:/Users/asger/OneDrive - Aarhus universitet/UNI/Machine Learning/Machine_Learning/Problemset 2/Data/campnet_attr.csv")

A <- as.matrix(campnet_Copy)
G <- graph.adjacency(A, mode = "undirected")

plot(G)

plot(G, layout = layout.fruchterman.reingold, vertex.label.cex = 0.75, vertex.label.color = "black", 
     vertex.size = 8, vertex.label.dist = 1, vertex.color = campnet_Attribute$Role,
     edge.arrow.size = 0.1)

hist(degree(G))

D <- diag(degree(G))

L <- D - A

if(colSums(A) == 0) {
  
  colSums(A) == 0
  
}


Laplacian_normalized <- function(A) {
  
  A <- A + t(A)
  G <- graph.adjacency(A)
  d <- degree(G)
  D <- diag(d)
  L <- D - A
  L_n <- t(d^(-1/2) * L) * d^(-1/2)
  
  L_n[is.na(L_n)] <- 0
  
  return(L_n)
  
}

```



```{r}

library(DescTools)

spec_clust <- function(ln, k) {
  
  ### Find the eigenvectors and eigenvalues of Lap
  
  eigen <- eigen(ln)
  length <- length(eigen$values)
  
  U <- matrix(eigen$vectors[,(length - k + 1):length], nrow = nrow(ln), ncol = k)
  X <- U / sqrt(rowSums(U^2))
  X[is.na(X)] <- 0
  
  Cluster <- kmeans(X, k, iter.max = 10, nstart = 1000)
  
  return((Cluster$cluster))
  
}

```


```{r}

cluster <- spec_clut(Ln, 3)

plot(G, layout = layout.fruchterman.reingold, vertex.label.cex = 0.75, vertex.label.color = "black", 
     vertex.size = 8, vertex.label.dist = 1, vertex.color = cluster,
     edge.arrow.size = 0.1)

```
Consider the S&P-100 data set from the slides and tutorials. Load in the data and estimate the network with λ = 0.6.

Calculate the density of the estimated network.

```{r}

sp100_dates <- read.csv("C:/Users/asger/OneDrive - Aarhus universitet/UNI/Machine Learning/Machine_Learning/Problemset 2/Data/sp100-dates - Copy.csv")
sp100_info <- read.csv("C:/Users/asger/OneDrive - Aarhus universitet/UNI/Machine Learning/Machine_Learning/Problemset 2/Data/sp100-info - Copy.csv")
sp100_returns <- read.csv("C:/Users/asger/OneDrive - Aarhus universitet/UNI/Machine Learning/Machine_Learning/Problemset 2/Data/sp100-returns - Copy.csv")

head(sp100_dates)
head(sp100_info)
head(sp100_returns)

library(space)

heatmap(cor(sp100_returns), Rowv = NA, Colv = NA, scale = "none")

Y <- matrix(0, nrow = nrow(sp100_returns), ncol = ncol(sp100_returns))

for (i in 1:ncol(sp100_returns)) {
  
  model <- lm(sp100_returns[,i] ~ rowMeans(sp100_returns))
  Y[,i] <- model$residuals
  
}

heatmap(cor(Y), Rowv = NA, Colv = NA, scale = "none")

sample <- space.joint(Y, lam1 = 0.6 * nrow(Y))
A <- 1*(sample$ParCor != 0) - diag(rep(1,ncol(Y)))

network <- graph.adjacency(A)

#### This is the density of the network ####

graph.density(network)

#### Calculate the normalized degree centrality, eigenvector centrality and Page Rank centrality of the vertices.

eigen_centrality(network)$vector

page_rank(network)$vector

degree(network) / (ncol(Y) - 1)

#### Is the graph connected?

plot(network)

#### well visually it is clearly not connected! But we can also make use of R for this.

A <- as.matrix(as_adjacency_matrix(network))

eigen(diag(degree(network)) - A)$vector[,ncol(Y)-1]
is_connected(network)

#### We also see that the eigen vector for the second last column is also 0 so it is not connected.

components(network)

#### Calculate the normalized Laplacian

ln <- Laplacian_normalized(A)

#### Apply spectral clustering to the network.

lo <- layout_with_kk(network)
lo <- norm_coords(lo, ymin=-1, ymax=1, xmin=-1, xmax=1)

plot(network, layout = lo*3, vertex.label.cex = 0.75, vertex.label.color = "black", 
     vertex.size = 5, vertex.label.dist = 1, vertex.label = sp100_info$Ticker.symbol, vertex.color = 1,
     edge.arrow.size = 0.1, rescale = FALSE)

cluster <- spec_clust(ln, 1)

plot(network, layout = lo*1.5, vertex.label.cex = 0.75, vertex.label.color = "black", 
     vertex.size = 7, vertex.label.dist = 1, vertex.label = sp100_info$Ticker.symbol, vertex.color = cluster,
     edge.arrow.size = 0.1, rescale = FALSE)

table(cluster)

```

```{r}

trump_adj <- as.matrix(read.csv("C:/Users/asger/OneDrive - Aarhus universitet/UNI/Machine Learning/Machine_Learning/Problemset 2/Data/trump_adj.csv"))
trump_names <- read.csv("C:/Users/asger/OneDrive - Aarhus universitet/UNI/Machine Learning/Machine_Learning/Problemset 2/Data/trump_names.csv")

head(trump_adj)
head(trump_names)

A <- trump_adj[,-1]

G <- graph.adjacency(A)

lo <- layout_with_kk(G)
lo <- norm_coords(lo, ymin=-1, ymax=1, xmin=-1, xmax=1)

plot(network, layout = lo, vertex.label.cex = 0.75, vertex.label.color = "black", 
     vertex.size = 7, vertex.label.dist = 1, vertex.label = trump_names$realDonaldTrump,
     edge.arrow.size = 0.1, rescale = FALSE)

graph.density(G)

A <- as.matrix(as_adjacency_matrix(G))

Lm <- Laplacian_normalized(A)

cluster <- spec_clust(Lm, 3)

plot(network, layout = lo, vertex.label.cex = 0.75, vertex.label.color = "black", 
     vertex.size = 7, vertex.label.dist = 1, vertex.label = trump_names$realDonaldTrump,
     vertex.color = cluster, edge.arrow.size = 0.1, rescale = FALSE)

```

```{r}

library(tidyverse)

got_s1_edges <- read.csv("C:/Users/asger/OneDrive - Aarhus universitet/UNI/Machine Learning/Machine_Learning/Problemset 2/Data/got-s1-edges.csv")

got_s1_nodes <- read.csv("C:/Users/asger/OneDrive - Aarhus universitet/UNI/Machine Learning/Machine_Learning/Problemset 2/Data/got-s1-nodes.csv")

View(got_s1_edges)
View(got_s1_nodes)

order <- match(got_s1_edges$Source, got_s1_nodes$Id)

got_s1_nodes$Community[order]

G <- graph.edgelist(as.matrix(got_s1_edges[,1:2]))

lo <- layoutByAttribute(G, attribute = 1)
lo <- norm_coords(lo, ymin=-1, ymax=1, xmin=-1, xmax=1)

plot(network, layout = lo*1, vertex.label.cex = 0.6, vertex.label.color = "black", 
     vertex.size = 7, vertex.label.dist = 1, vertex.label = got_s1_nodes$Label[order], edge.arrow.size = 0.1, rescale = TRUE, vertex.color = got_s1_nodes$Community[order])

```

```{r}

ho_pe <- read.csv("C:/Users/asger/OneDrive - Aarhus universitet/UNI/Machine Learning/Machine_Learning/Problemset 2/Data/ho-pe.csv", row.names = 1)
ho_states <- read.csv("C:/Users/asger/OneDrive - Aarhus universitet/UNI/Machine Learning/Machine_Learning/Problemset 2/Data/ho-states.csv")

head(ho_pe)
head(ho_states)

heatmap(cor(ho_pe))

f_hat <- as.matrix(ho_pe) %*% eigen(cor(ho_pe))$vector[,1]

Y <- matrix(0, nrow = nrow(ho_pe), ncol(ho_pe))

for (i in 1:ncol(ho_pe)) {
  
  model <- lm(ho_pe[,i] ~ f_hat)
  Y[,i] <- model$residuals
  
}

heatmap(cor(Y))

```
It is much less red!

```{r}

library(space)

lambda.range <- seq(0.1, 3.0, 0.1)

bic.vec  <- rep(0, length(lambda.range))
npar.vec <- rep(0, length(lambda.range))

P.hat.list  <- list()

network.trace <- matrix(0, length(lambda.range), ncol(Y) * (ncol(Y) - 1) / 2 )

for( l in 1:length(lambda.range) ) {
  
  # Run SPACE for the candidate lambda value times T:
  results <- hush(space.joint(Y, lam1 = lambda.range[l] * nrow(Y)))
  
  # The hush() function silences annoying print messages.
  
  # Calculate the BIC and number of non-zero partial correlations:
  bic_res <- bic_space(results)
  bic.vec[l] <- bic_res[[1]]
  npar.vec[l] <- bic_res[[2]]
  
  # Save the partial correlation matrix into our list:
  P.hat   <- results$ParCor 
  P.hat.list[[l]] <- P.hat
  
  # Save the lower triangular part of the partial correlation matrix as a row in the network.trace matrix:
  network.trace[l,] <- (P.hat[ lower.tri(P.hat)==1 ])[1:(ncol(Y) * (ncol(Y)-1) / 2)]
  
  # Print message every 5 lambda values:
  if( l %% 5 == 0){
    cat("Done with", l, "/", length(lambda.range), "lambda values.\r")
  }
}

# The index of the optimal lambda:
lambda.opt.ind <- which(bic.vec == min(bic.vec))

# The selected lambda: 
lambda.range[lambda.opt.ind]

# The selected partial correlation matrix:
P.hat <- P.hat.list[[lambda.opt.ind]]

network <- hush(space.joint(Y, lambda.range[lambda.opt.ind] * nrow(Y)))

```

```{r}

library(igraph)

A <- 1*(network$ParCor != 0) - diag(rep(1,ncol(Y)))

Ag <- graph.adjacency(as.matrix(A), "undirected")

graph.density(Ag)

hist(degree(Ag))
table(degree(Ag))

```
We clearly see a power law structure in this network, many people with few connections and a few people with many connections!

```{r}

Ln <- Laplacian_normalized(A)

cluster <- spec_clust(Ln, 3)

plot(Ag, vertex.color = cluster)

### install.packages("mosaic")
library(mosaic)

df <- cbind(ho_states,cluster)

mUSMap(data = df, key = "State", fill = "cluster")

```
Maybe some political map of the US

```{r}

Blockbuster <- function(Y,k) {
  
  cov_matrix <- cor(Y)
  U <- eigen(cov_matrix)$vector[,1:k]
  euclidean_norm <- sqrt(rowSums(U)^2)
  
  X <- U / euclidean_norm
  
  cluster <- kmeans(X, k, iter.max = 10, nstart = 1000)
  
  return(cluster$cluster)
  
}

cluster2 <- Blockbuster(Y,3)

df <- cbind(df, cluster2)

mUSMap(data = df, key = "State", fill = "cluster2")

```


```{r}

### install.packages("neuralnet")
library(neuralnet)

Hitters <- read.csv("C:/Users/asger/OneDrive - Aarhus universitet/UNI/Machine Learning/Machine_Learning/Problemset 2/Data/Hitters - Copy.csv")

Hitters_int <- Hitters %>%
  select(-League, -Division, -NewLeague, -X)

summary(Hitters_int)

hist(Hitters_int$Salary)

library(mice)

imp <- mice(Hitters_int)
Hitters_full <- complete(imp)

summary(Hitters_full)

##### Normalisation:
max_vec <- apply(Hitters_full, 2, max)
min_vec <- apply(Hitters_full, 2, min)
Hitters_full_nom <- as.data.frame(scale(Hitters_full, center = min_vec, scale = max_vec - min_vec))

### Split data into training and test

train_ind <- sample(1:length(Hitters_full_nom), length(Hitters_full_nom) * 0.7)

train_data <- Hitters_full_nom[train_ind,]
test_data <- Hitters_full_nom[-train_ind,]

#### Train linear model

library(forecast)

lm_model <- lm(log(Salary) ~ ., data = train_data)
pred_OLS <- predict(lm_model, test_data)

MSE_OLS <- mean((pred_OLS - test_data$Salary)^2)

accuracy(pred_OLS, test_data$Salary)

#### Train Lasso model

library(glmnet)

X <- train_data %>%
  select(-Salary) %>%
  as.matrix(.)
Y <- train_data$Salary

cv_out <- cv.glmnet(X, log(Y), alpha = 1, nfolds = 100)
plot(cv_out)

lambda_min <- cv_out$lambda.min

lasso_model <- glmnet(X, log(Y), alpha = 1, lambda = lamda_min)

coef(lasso_model)

pred_lasso <- as.vector(predict(lasso_model, newx = as.matrix(test_data[,-17])))

accuracy(pred_lasso, test_data$Salary)

MSE_lasso <- mean((pred_lasso - test_data$Salary)^2)

#### Train neural network model

nn_model <- neuralnet(log(Salary) ~ ., data = train_data, hidden = c(5,3,2), linear.output = T)
plot(nn_model)
pred_nn <- predict(nn_model, test_data)

MSE_nn <- mean((pred_nn - test_data$Salary)^2)

#### Print models

print(c(MSE_OLS, MSE_nn, MSE_lasso))

plot(test_data$Salary, pred_nn, col = "red", main = "Neural Network")
abline(0, 1)

plot(test_data$Salary, pred_OLS, col = "blue", main = "Linear Regression")
abline(0,1)

plot(test_data$Salary, pred_nn, col = "red", main = "OLS and NN", xlim = c(0,6), ylim = c(0,1))
points(test_data$Salary, pred_OLS, col = "blue")
abline(0,1)

```

```{r}

plot(test_data$Salary, pred_lasso)
abline(0,1)

```

